---
title: "HM5"
author: "狄豪飛"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE, include=FALSE}
# Cargar las librerías necesarias
library(tidyverse)
```

## Student Stress Factors: A Comprehensive Analysis
It contains various features that impact a student's stress level, grouped into Psychological, Physiological, Environmental, Academic, and Social factors.


### Understand the Dataset and Objectives

#### Dataset Overview:

Psychological Factors: 
anxiety_level, self_esteem, mental_health_history, depression

Physiological Factors:
headache, blood_pressure, sleep_quality, breathing_problem

Environmental Factors: 
noise_level, living_conditions, safety, basic_needs

Academic Factors:
academic_performance, study_load, teacher_student_relationship, future_career_concerns

Social Factor: 
social_support, peer_pressure, extracurricular_activities, bullying

Objectives:
To understand which features have the most impact on a student's stress level.

```{r, message = FALSE}
# Cargar el conjunto de datos
df <- read.csv("C:/Users/georg/OneDrive/Desktop/Data Science/HM5/HM5/StressLevelDataset.csv")

# Visualizar las primeras filas del conjunto de datos
head(df)

# Obtener un resumen estadístico
summary(df)

# Verificar los tipos de datos de cada columna
str(df)

# Comprobar valores faltantes
missing_values <- sapply(df, function(x) sum(is.na(x)))
print(paste("Valores faltantes por columna:"))
print(missing_values)
```



### Data Cleaning

```{r}
# Verificar registros duplicados
duplicated_rows <- duplicated(df)
print(paste("Número de registros duplicados:", sum(duplicated_rows)))

# Eliminar registros duplicados si los hay
if (sum(duplicated_rows) > 0) {
  df <- df[!duplicated_rows, ]
}

# Función para identificar valores atípicos
identify_outliers <- function(column, data) {
  q1 <- quantile(data[[column]], 0.25)
  q3 <- quantile(data[[column]], 0.75)
  iqr <- q3 - q1
  
  lower_bound <- q1 - (1.5 * iqr)
  upper_bound <- q3 + (1.5 * iqr)
  
  outliers <- subset(data, data[[column]] < lower_bound | data[[column]] > upper_bound)
  
  return(nrow(outliers))
}

# Características seleccionadas para examinar valores atípicos
selected_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance')

# Identificar y contar valores atípicos para las características seleccionadas
outliers_count <- sapply(selected_features, identify_outliers, data=df)
print("Número de valores atípicos por columna:")
print(outliers_count)

```

#### Missing Values:
There are no missing values in any of the columns.

#### Duplicate Rows:
There are no duplicate rows in the dataset.

Since the dataset doesn't have any missing values or duplicates, we can proceed to the next step without needing to clean the data in these aspects.



### Data Transformation

```{r}

# Resumen estadístico para entender la distribución de cada característica
summary(df)

# Escalado Min-Max
min_vals <- apply(df, 2, min, na.rm = TRUE)
max_vals <- apply(df, 2, max, na.rm = TRUE)
df_minmax <- as.data.frame(scale(df, center = min_vals, scale = max_vals - min_vals))

# Normalización Z-score
df_zscore <- as.data.frame(scale(df))

# Visualizar las primeras filas de los conjuntos de datos escalados para comparar
head(df_minmax)
head(df_zscore)


```

### The statistical summary provides the following insights:

#### Range of Features:
Most features seem to be scored on a scale that varies from feature to feature. For instance, anxiety_level ranges from 0 to 21, while blood_pressure ranges from 1 to 3.

#### Distribution: 
The mean and standard deviation for each feature provide an idea of the data distribution. For example, self_esteem has a mean of approximately 17.78 and a standard deviation of approximately 8.94, indicating a fairly wide distribution.

#### Outliers: 
There don't appear to be any extreme values that could be considered outliers based on the min-max range and quartile values.



### Univariate Analysis
```{r}

# Función para trazar análisis univariante para una columna dada
plot_univariate_analysis <- function(column, data, scaled_data1, scaled_data2) {
  
  # Crear una ventana gráfica con 3 paneles
  par(mfrow=c(1, 3))
  
  # Histograma para datos originales
  hist(data[[column]], main=paste("Original: ", column), xlab=column, col="blue", breaks=20, freq=FALSE)
  
  # Histograma para datos escalados por Min-Max
  hist(scaled_data1[[column]], main=paste("Min-Max Scaled: ", column), xlab=column, col="green", breaks=20, freq=FALSE)
  
  # Histograma para datos normalizados por Z-score
  hist(scaled_data2[[column]], main=paste("Z-Score Normalized: ", column), xlab=column, col="red", breaks=20, freq=FALSE)
}

# Características seleccionadas para demostrar el enfoque
selected_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance')

# Generar gráficos para las características seleccionadas
for (feature in selected_features) {
  plot_univariate_analysis(feature, df, df_minmax, df_zscore)
}
```

#### Observations:

#### Anxiety Level: 
Appears to be slightly left-skewed in all forms. This suggests that most students have moderate to high anxiety levels.

#### Self Esteem: 
This feature shows a nearly normal distribution in its original and scaled forms, indicating a balanced range of self-esteem levels among students.

#### Blood Pressure: 
The distribution looks bimodal, with two peaks. This could indicate two different groups within the data, perhaps separated by some other feature or condition.

#### Academic Performance: 
Shows a nearly uniform distribution in its original form but becomes a bit more bell-shaped when scaled.

These visualizations give us valuable insights into each feature's distribution, which can be useful for feature selection and model building later on. It also helps in understanding how scaling methods affect the distribution.



### Bivariate and Multivariate Analysis
```{r}

# Cargar la biblioteca para gráficos de correlación
library(corrplot)

# Calcular la matriz de correlación
cor_matrix <- cor(df, use = "complete.obs")

# Dibujar la matriz de correlación
corrplot(cor_matrix, method = "color")


# Función corregida para trazar gráficos de caja con colores
plot_bivariate_analysis <- function(target, features, data) {
  for (feature in features) {
    p <- ggplot(data, aes(x=factor(get(target)), y=get(feature), fill=factor(get(target)))) +
      geom_boxplot() +
      scale_fill_brewer(palette="Set1") +
      ggtitle(paste(feature, " vs ", target)) +
      xlab(target) +
      ylab(feature)
    print(p)
  }
}

# Características seleccionadas
selected_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance', 
                       'headache', 'sleep_quality', 'noise_level', 'living_conditions', 
                       'study_load', 'teacher_student_relationship', 
                       'social_support', 'peer_pressure', 'extracurricular_activities', 'bullying')

# Generar gráficos de caja con colores
plot_bivariate_analysis('stress_level', selected_features, df)



# Función para realizar pruebas ANOVA
perform_anova_tests <- function(target, features, data) {
  results <- data.frame(Feature=character(), P_value=numeric())
  for (feature in features) {
    anova_result <- aov(as.formula(paste(target, "~", feature)), data=data)
    p_value <- summary(anova_result)[[1]][["Pr(>F)"]][1]
    results <- rbind(results, data.frame(Feature=feature, P_value=p_value))
  }
  return(results)
}

# Realizar pruebas ANOVA
anova_results <- perform_anova_tests('stress_level', selected_features, df)

# Imprimir los resultados
print(anova_results)

```

#### Bivariate and Multivariate Analysis

##### Correlation Matrix:
The heatmap shows the correlation between all the features and the target variable (stress_level). A higher absolute value indicates a stronger relationship.

Features like anxiety_level, depression, and mental_health_history seem to have a relatively higher correlation with stress_level.

##### Bivariate Analysis:

Box plots for the target variable (stress_level) against selected features were plotted.

Anxiety Level: Higher levels of stress appear to be associated with higher anxiety levels.

Self Esteem: Lower stress levels seem to correspond with higher self-esteem.

Blood Pressure: There doesn't appear to be a clear trend between blood pressure and stress levels, indicating a weaker relationship.

Academic Performance: Similarly, academic performance also doesn't show a strong trend with respect to stress levels.

These analyses can guide feature selection and help in understanding how different features interact with the target variable.



### Statistical Testing

```{r}
# Función para realizar pruebas t de Student o ANOVA
perform_stat_tests <- function(target, features, data) {
  results <- data.frame(Feature=character(), P_value=numeric())
  
  for (feature in features) {
    if (length(unique(data[[target]])) == 2) {
      t_result <- t.test(get(feature) ~ get(target), data=data)
      p_value <- t_result$p.value
    } else {
      anova_result <- aov(as.formula(paste(feature, "~", target)), data=data)
      p_value <- summary(anova_result)[[1]][["Pr(>F)"]][1]
    }
    
    results <- rbind(results, data.frame(Feature=feature, P_value=p_value))
  }
  
  return(results)
}

# Características seleccionadas para pruebas estadísticas
selected_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance',
                       'headache', 'sleep_quality', 'noise_level', 'living_conditions',
                       'study_load', 'teacher_student_relationship',
                       'social_support', 'peer_pressure', 'extracurricular_activities', 'bullying')

# Realizar pruebas t de Student o ANOVA
stat_results <- perform_stat_tests('stress_level', selected_features, df)

# Imprimir los resultados
print("Resultados de las pruebas estadísticas:")
print(stat_results)

```

#### Observations:
All p-values are very close to zero, which suggests that there is a statistically significant difference in these features across different levels of stress.

This confirms the observations made during the bivariate analysis and adds statistical evidence to the relationships.



### Data Visualization - Summary Plots

```{r}
# Función para hacer gráficos de barras para variables categóricas
plot_categorical <- function(column, data) {
  p <- ggplot(data, aes_string(x=column)) +
    geom_bar(aes(fill=column), position="dodge") +
    ggtitle(paste("Distribución de", column))
  print(p)
}

# Características categóricas seleccionadas para visualización
categorical_features <- c('mental_health_history', 'living_conditions', 'teacher_student_relationship', 'bullying')

# Dibujar gráficos de barras
for (feature in categorical_features) {
  plot_categorical(feature, df)
}

# Función para hacer gráficos de densidad para variables numéricas
plot_density <- function(column, data) {
  p <- ggplot(data, aes_string(x=column)) +
    geom_density(aes(fill=column), alpha=0.5) +
    ggtitle(paste("Densidad de", column))
  print(p)
}

# Características numéricas seleccionadas para visualización
numeric_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance')

# Dibujar gráficos de densidad
for (feature in numeric_features) {
  plot_density(feature, df)
}


# Función para crear gráficos de violín para una columna dada contra la variable objetivo
plot_violin <- function(target, features, data) {
  for (feature in features) {
    p <- ggplot(data, aes_string(x=target, y=feature, fill=target)) +
      geom_violin(scale="area") +
      ggtitle(paste(feature, " vs ", target)) +
      xlab(target) +
      ylab(feature)
    print(p)
  }
}

# Características numéricas seleccionadas para visualización con gráficos de violín
selected_features <- c('anxiety_level', 'self_esteem', 'blood_pressure', 'academic_performance')

# Crear gráficos de violín para las características seleccionadas contra 'stress_level'
plot_violin('stress_level', selected_features, df)


```

### Observations:

#### Anxiety Level: 
The plot reiterates that higher levels of stress are associated with higher anxiety levels.

#### Self Esteem:
Lower levels of stress seem to correspond with higher self-esteem, confirming our previous observations.

#### Blood Pressure:
The plot shows a wide range of blood pressure levels across different stress levels, but there isn't a clear trend.

#### Academic Performance: 
Similar to blood pressure, there isn't a strong trend between academic performance and stress levels.

These summary plots encapsulate the key findings from the EDA and provide a visual summary that can be useful for presentations or reports.



#### Summary and Insights

##### - Most features exhibit significant relationships with the stress_level based on statistical testing.

#####  - Features like anxiety_level, self_esteem, and depression appear to have a stronger influence on stress_level.

##### - The dataset did not require much cleaning, but feature scaling was essential for understanding how different features behave under normalization.
